---
title: "LearningMovement"
author: "Santiago Patiño-Giraldo"
date: "16/8/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Learning Machine

The aim of this project is create a model to predict the class of activity a person made based on variables taken by accelerometers on the belt, forearm, arm, and dumbell of 6 participants

## Loading and preparing data

```{r, echo=FALSE,eval=FALSE}
download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile = "training.csv",method = "curl")

download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile = "testing.csv",method = "curl")
```

Create two dataframes

```{r}
training <- read.csv("training.csv")
testing <- read.csv("testing.csv")

library(caret)
library(lubridate)
library(rpart)
```
Looking for the data

```{r}
str(training)
```

Some variables needs transformation, is presented as factors but there are numeric
values

```{r}
training2 <- training
testing2 <- testing
training2$cvtd_timestamp <- dmy_hm(training2$cvtd_timestamp)
testing2$cvtd_timestamp <- dmy_hm(testing2$cvtd_timestamp)

tonumeric <- function(x){
        for(i in 7:159){
                if(is.factor(x[,i]) | is.logical(x[,i])){
                        x[,i] <- as.numeric(x[,i])
                }
        }
        x
}
training2 <- tonumeric(training2)
testing2 <- tonumeric(testing2)
```

The training consist in 19622 observations of 160 variables. I discarded the measures
with more than 50% of NA's

```{r}
countna <- function (x){
        w <- vector()
        for(i in 1:160){
        y <- sum(is.na(x[i]))/nrow(x)
        
        if(y >= 0.5){
        w[i] <- FALSE        
        }
        else{
                w[i] <- TRUE
        }
        }
        w
}

excluding <- countna(training2)
training2 <- training2[excluding]
testing2 <- testing2[excluding]
str(training2)
```

I will probe 3 different prediction models: random forest, decision tree and bgm
I cluster in 6 the training data to do that.

```{r}
set.seed(19862)
folds.test <- createFolds(y=training2$classe,k=6,list=TRUE)
```

## Decision tree
```{r}
model1 <- rpart(classe~.,data=training2[folds.test$Fold1,],method="class")

acc_prediction <- function(model,matrix,base,clase=NULL){
        y <- vector()
        for(i in 1:6){
                if(i==base){
                        
                }
                else{
                       
                        if(is.null(clase)){
                         x <- predict(model,matrix[folds.test[[i]],])
                        }
                        else{
                         x <- predict(model,matrix[folds.test[[i]],],type="class")       
                        }
                       y[i] <- confusionMatrix(x,
                                matrix[folds.test[[i]],93])$overall[1]         
                }
        }
        y
}


acc_model1 <- acc_prediction(model1,training2,1,"Y")

```

## Random Forest

```{r}
model2 <- train(classe~.,data=training2[folds.test$Fold2,],method="rpart")
acc_model2 <- acc_prediction(model2,training2,2)
```

## Generalized Boosted Regression

```{r, echo=FALSE}
model3 <- train(classe~.,data=training2[folds.test$Fold3,],method="gbm")
```

```{r}
acc_model3 <- acc_prediction(model3,training2,3)
```

Showing accuracies

```{r}
acc_model1
acc_model2
acc_model3
```

Models 1 and 3 could be overfitted but model 2 has low accuracy. 
Model3 takes a lot of time for training
I chose model 1

```{r}
acc_testing <- predict(model1,testing2)
acc_testing
```